---
title: "Statistical Inference course project"
author: "Dmitri Perov"
date: "December 26, 2015"
output: html_document
---

Overview 
--------

This project aims to investigate the exponential distribution in R and compare it with the Central Limit Theorem. The sample mean is  compared  to the theoretical mean of the distribution. The sample variance  is compared  to the theoretical variance of the distribution. It is shown then that the distribution is approximately normal.


Simulations
-----------

Get the means of the 1000 simulations
```{r setup}
set.seed(1)
n <- 1000        ## no of runs
sample.size <- 40 ## 40 samples in each run
lambda <- 0.2   ##variable input
dist <- matrix(rexp(sample.size*n, rate=lambda), ncol = sample.size, nrow=n)
dist.means <- rowMeans(dist) ##a  vector of n length with averages in each row based on sample size means
```




Sample Mean versus Theoretical Mean
-----------------------------------

```{r Q1, echo=FALSE}
hist(dist.means,breaks=sample.size,prob=T,col="green",main="Density of means",ylab="density")
abline(v=mean(dist.means),col="black",lwd=4)

center.ac <- mean(dist.means)
center.th <- 1/lambda
```

The Theoretical center of the distribution is calculated as $1/\lambda$ = 1/0.2 = `r center.th`. The center of the distribution is `r center.ac`. The black line in the above plot displays the center.



Sample Variance versus Theoretical Variance
-------------------------------------------

```{r Q2}
sd.ac <- sd(dist.means)
sd.th <- (1/lambda)*(1/sqrt(sample.size))

var.ac <- sd.ac^2 ## = var(dist.means)
var.th <-((1/lambda)*(1/sqrt(sample.size)))^2
```

Standard Deviation of the distribution is `r sd.ac` with the theoretical SD calculated as `r sd.th`.
The Theoretical variance is calculated as $(\frac{1}{\lambda} * \frac{1}{\sqrt{n}})^2$ = `r var.th`. Actual variance of the distribution is `r var.ac`

Distribution
------------

```{r Q3, echo=FALSE}
xfit <- seq(min(dist.means), max(dist.means), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(sample.size)))
hist(dist.means,breaks=sample.size,prob=T,col="green",main="Density of means",ylab="density")
lines(xfit, yfit, pch=22, col="black", lty=5)
```


```{r Q3_2, echo=FALSE}
qqnorm(dist.means,col="red" ) #normal QQ plot
qqline(dist.means,lwd=3) ##theoretical line
```

In the first plot we have overlayed a normal distribution (in black) over the density plot taken from the means of the exponential distribution. To confirm the distribution we use a qqnorm plot and overlay the theoretical line.
We notice in the QQ plot that most of the red is on the theoretical normal line and only deviates at the begninning and end due to the skewness of the exponential distribution.


